---
title: "Figures_in_section_4_2"
output:
  html_document:
    code_folding: hide
date: "2026-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries:
```{r message=FALSE, warning=FALSE}
library(Rcpp)
library(dplyr)
library(ggplot2)
library(tibble)
library(tidyr)
library(tidyverse)
library(latex2exp)
library(ggrepel)
library(grid)
library(cowplot)
library(extrafont)
```

Sourced files:
```{r message=FALSE, warning=FALSE}
# helper functions for simulation and plotting
source('Brownian_motion.R',keep.source = TRUE)
source('EM.R',keep.source = TRUE)
source('plots.R',keep.source = TRUE)

sourceCpp("DoubleWell.cpp") # functions for Strang splitting and likelihood 
sourceCpp("b_vec_selection.cpp") # functions for Strang splitting and likelihood 
```

## Figure 1
```{r message=FALSE, warning=FALSE}
eps1 <- 0.1
eps2 <- 0.25
y1 <- -0.05
y2 <- 0
sigma1 <- 1.7
sigma2 <- 1.2
F1 <- function(x){(x-x^3-y1)/eps1}
F2 <- function(x){(x-x^3-y2)/eps2}
G1 <- function(x){sigma1}
G2 <- function(x){sigma2}

x0 <- c(0)
n_paths <- 1
h <- 0.0001
T <- 30
m <- T/h +1
times <- seq(0,T,by=h)

W_increments <- rnorm(m,0,sqrt(h))
W_increments[1]<-0

EM <-  function(F,G){
  X_e <- rep(x0,m)
  for(i in (2:m)){
    X_e[i] <- X_e[i-1] + F(X_e[i-1])*h + G(X_e[i-1])*W_increments[i] } 
  return(X_e)
}
path1 <- EM(F1,G1)
path2 <- EM(F2,G2)

samplepath_p <- plot_doublewell_sample_paths_with_same_noise(path1,path2)
samplepath_p
ggsave(plot = samplepath_p, width = 5.5, height = 4.5, dpi = 300, filename = "samplepaths 220726.pdf")
xs <- seq(-1.5,1.5,by=0.001)
U1 <- function(x){(0.25*x^4 - 0.5* x^2 + x*y1)/eps1}
U2 <- function(x){(0.25*x^4 - 0.5* x^2 + x*y2)/eps2}
U1s <- U1(xs)
U2s <- U2(xs)
data1 <- data.frame(cbind(xs,U1s))
data2 <- data.frame(cbind(xs,U2s))


potential_p <-  plot_doublewell_potential(data1,data2)
potential_p
ggsave(plot = potential_p, width = 4, height = 4.5, dpi = 300, filename = "potentials 220726.png")
```




## Figure 2
```{r message=FALSE, warning=FALSE}

#Number of simulations:
M <- 2000
#Parameters:
y <- -0.05
eps <- 0.1
sigma <- 1.7
#Choice of transition density:
x0 <- -1.3
x0_left <- x0
h <- 0.08

#Splitting scheme functions:
F <- function(x){(x-x^3-y)/eps}
G <- function(x){sigma}
DF <- function(x){(1-3*x^2)/eps}
A <- function(c){DF(c)}
btilde <- function(c){ c - F(c)/DF(c) }
N <- function(x,c){(-x^3+3*c^2*x-2*c^3)/eps}
muh <- function(x,h,c){exp(A(c)*h)*x+(1-exp(A(c)*h))*btilde(c)}
Omegah <- function(h,c){  sigma^2 * (exp(2*A(c)*h)-1)/(2*A(c))  }
fh_fun <- function(x, h,c) {
  ODEfun <- function(t, state, parms) {
    u1 <- state[1]
    du1 <- N(u1,c)
    list(c(du1)) }
  out <- deSolve::ode(y = x, times = c(0, h), func = ODEfun, parms = NULL, method = 'lsoda')
  as.numeric(out[nrow(out), -1])
}

#Simulating EM steps:
hEM <- h/2000
EMtimes <- seq(0,h,by=hEM)
EMsim <- rep(NA,M)
for(i in 1:M){
  EMsteps <-  EM_sim_fast(F,G,x0,EMtimes,n_paths=1,progress = FALSE)[[1]]
  EMsim[i] <- EMsteps[nrow(EMsteps),1]
}
EMdata <- data.frame(EMsim)
EMdata$scheme <- "True distribution (EM)"
EMdata$x0 <- x0
names(EMdata) <- c("sim", "scheme", "Initial value")

#Function for simulating one-step predictions:
StrangSim <- function(x0,h,c,M){
  X0 <- rep(x0,M)
  xi <- rnorm(M,0,sqrt(Omegah(h,c)))
  step1 <- mapply(fh_fun,X0,h/2,c)
  step2 <- mapply(muh,step1,h,c)+xi
  step3 <- mapply(fh_fun,step2,h/2,c)
  return(step3)
}

#Splitting around negative stable fixed point:
StrangDataFix <- data.frame(StrangSim(x0,h, min(Re(polyroot(c(-y,1,0,-1))))  ,M))
StrangDataFix$scheme <- "Fixed point"
StrangDataFix$x0 <- x0
names(StrangDataFix) <- c("sim","scheme", "Initial value")

#Splitting around opposite stable fixed point:
StrangDataOpp <- data.frame(StrangSim(x0,h, max(Re(polyroot(c(-y,1,0,-1))))  ,M))
StrangDataOpp$scheme <- "Opposite fixed point"
StrangDataOpp$x0 <- x0
names(StrangDataOpp) <- c("sim","scheme","Initial value")

#Local splitting around x0:
StrangDataEachPoint <- data.frame(StrangSim(x0,h, x0  ,M))
StrangDataEachPoint$scheme <- "Local Strang"
StrangDataEachPoint$x0 <- x0
names(StrangDataEachPoint) <- c("sim","scheme","Initial value")

#Combine data for x0 in A1
data <- rbind(EMdata, StrangDataFix, StrangDataOpp, StrangDataEachPoint)

#Repeat for x0 in A3
x0 <- 0.6
EMsim2 <- rep(NA,M)
for(i in 1:M){
  EMsteps2 <-  EM_sim_fast(F,G,x0,EMtimes,n_paths=1,progress = FALSE)[[1]]
  EMsim2[i] <- EMsteps2[nrow(EMsteps2),1]
}
EMdata2 <- data.frame(EMsim2)
EMdata2$scheme <- "True distribution (EM)"
EMdata2$x0 <- x0
names(EMdata2) <- c("sim", "scheme", "Initial value")
StrangDataFix2 <- data.frame(StrangSim(x0,h, max(Re(polyroot(c(-y,1,0,-1))))  ,M))
StrangDataFix2$scheme <- "Fixed point"
StrangDataFix2$x0 <- x0
names(StrangDataFix2) <- c("sim","scheme", "Initial value")
StrangDataOpp2 <- data.frame(StrangSim(x0,h, min(Re(polyroot(c(-y,1,0,-1))))  ,M))
StrangDataOpp2$scheme <- "Opposite fixed point"
StrangDataOpp2$x0 <- x0
names(StrangDataOpp2) <- c("sim","scheme","Initial value")
StrangDataEachPoint2 <- data.frame(StrangSim(x0,h, x0  ,M))
StrangDataEachPoint2$scheme <- "Local Strang"
StrangDataEachPoint2$x0 <- x0
names(StrangDataEachPoint2) <- c("sim","scheme","Initial value")

data2 <- rbind(EMdata2,StrangDataFix2, StrangDataOpp2,StrangDataEachPoint2)
data <- rbind(data, data2)

p <- plot_doublewell_transition_densities(data)
p
```

## Figure 3(a)
```{r fig.height=8, fig.width=16, message=FALSE, warning=FALSE}
y <- -0.05
eps <- 0.1
sigma <- 1.7

Bias_fun <- function(c, x,abs=FALSE){
  denom <- 1/(24*eps^3)
  top <- (-6*c^3+3*y)*x^4 +
    x^3*(-36*c^4-24*sigma^2*eps+36*c^2-8) +
    x^2*(54*(c^3-y)*(c^2-1/3)) +
    x*(-24*c^6+54*c^2*eps*sigma^2+36*c^3*y-12*sigma^2*eps-12*y^2) +
    6*c^5-9*c^4*y+c^3*(-18*sigma^2*eps-4) +
    6*c^2*y+12*eps*sigma^2*y
  if (abs){
    return(abs(top*denom))
  }else{
    return(top*denom)
  }
}

# Function definition
DVar_fun <- function(c, x,abs=FALSE){
  Nx  <- (-x^3 + 3*c^2*x - 2*c^3) / eps
  dNx <- (-3*x^2 + 3*c^2) / eps
  d2Nx <- -6*x / eps
  d3Nx <- -6 / eps
  
  A <- (1 - 3*c^2) / eps
  b <- c - (c - c^3 - y) / (1 - 3*c^2)
  
  Dvar <- (-1/12)*sigma^2*d2Nx*Nx -
    (1/3)*sigma^2*d2Nx*A*(x-b) +
    (1/6)*sigma^2*dNx^2 +
    (1/3)*sigma^2*A*dNx -
    (1/6)*sigma^4*d3Nx
  
  if (abs){
    return(abs(Dvar))
  }else{
    return(Dvar)
  }
}

x0_vals <- c(-1, -0.1, 1.05)

c_grid <- seq(-1.5, 1.5, length.out = 3001)

df_all <- bind_rows(lapply(x0_vals, function(x0) {

  Bvals <- sapply(c_grid, function(c) Bias_fun(c, x0,abs=FALSE))
  Vvals <- sapply(c_grid, function(c) DVar_fun(c, x0,abs=FALSE))

  data.frame(
    c = c_grid,
    absB = as.numeric(Bvals),
    absVar = as.numeric(Vvals),
    x0 = x0
  )
}))

df_all <- df_all %>%
  group_by(x0) %>%
  mutate(max_regret = pmax(abs(absB), sqrt(abs(absVar)))) %>%
  #mutate(max_regret = absB^2+ abs(absVar)) %>%
  ungroup()
df_minimax <- df_all %>%
  group_by(x0) %>%
  slice_min(order_by = max_regret, n = 1, with_ties = FALSE) %>%
  ungroup()

tol <- 1

lex <- df_all %>%
  group_by(x0) %>%
  group_map(~{
    df <- .x
    x0_val <- .y$x0[[1]]

    # V -> B
    Vmin <- min(abs(df$absVar))
    cand_V <- df[abs(df$absVar) <= Vmin + tol, ]

    pt_V <- cand_V[which.min(abs(cand_V$absB)), ]
    

    # B -> V
    Bmin <- min(abs(df$absB))
    cand_B <- df[abs(df$absB) <= Bmin + tol, ]
    
    pt_B <- cand_B[which.min(abs(cand_B$absVar)), ]
    

    data.frame(
      x0 = x0_val,
      V_then_B_c    = pt_V$c,
      V_then_B_absB = pt_V$absB,
      V_then_B_absVar = pt_V$absVar,
      B_then_V_c    = pt_B$c,
      B_then_V_absB = pt_B$absB,
      B_then_V_absVar = pt_B$absVar
    )
  }) %>%
  bind_rows()

fixed_df <- bind_rows(lapply(x0_vals, function(x0) {

  roots <- polyroot(c(-y, 1, 0, -1))
  roots <- Re(roots[abs(Im(roots)) < 1e-8])

  if (length(roots) == 0) {
    data.frame(
      x0 = numeric(0),
      c = numeric(0),
      absB = numeric(0),
      absVar = numeric(0),
      type = character(0)
    )
  } else {
    data.frame(
      x0 = x0,
      c = roots,
      absB = sapply(roots, function(c) Bias_fun(c, x0,abs=FALSE)),
      absVar = sapply(roots, function(c) DVar_fun(c, x0,abs=FALSE)),
      type = "FixedPoint"
    )
  }
}))
legend_labels <- c(
  "Fixed points",
  "Minimizer of Local Variance (MoLV)",
  "Minimizer of Local Bias (MoLB)"
)


special_df <- bind_rows(
  # df_minimax %>%
  #   transmute(x0, c, absB, absVar, type = "Argmin Max{|B|,sqrt(|V|)}"),

  lex %>%
    transmute(
      x0,
      c = V_then_B_c,
      absB = V_then_B_absB,
      absVar = V_then_B_absVar,
      type = "Argmin|B| subj. to Argmin|V|"
    ),

  lex %>%
    transmute(
      x0,
      c = B_then_V_c,
      absB = B_then_V_absB,
      absVar = B_then_V_absVar,
      type = "Argmin|V| subj. to Argmin|B|"
    ),

  fixed_df %>%
    dplyr::select(x0, c, absB, absVar, type)
)
special_df <- special_df %>%
  mutate(
    nudge_x = ifelse(x0>1,150-absB,ifelse(x0 > 0, 120 - absB, -160 - absB))
  )

# order of facets
df_all$x0 <- factor(df_all$x0, levels = c(-1, -0.1, 1.05))

p1 <- plot_local_bias_and_var(
  df_all = df_all,
  special_df = special_df,
  legend_labels = legend_labels
)

p1
#ggsave(plot=p1,width=16,height=8,dpi=300,filename='BiasVarianceTradeoff.png')

```

## Figure 3(b)
```{r fig.height=10, fig.width=16, message=FALSE, warning=FALSE}
inv <- function(x,y,sigma,eps){exp(2*(0.5*x^2-0.25*x^4-y*x)/(sigma^2*eps))}
avg_absbias <- function(c,which_well){
  
  if (which_well =='left'){
    range <- c(-Inf,-1/sqrt(3))
  }else if (which_well == 'right'){
    range <- c(1/sqrt(3),Inf)
  }else if (which_well == 'middle'){
    range <- c(-1/sqrt(3),1/sqrt(3))
  }
  Z <- integrate(inv,-Inf,Inf,y=y,sigma=sigma,eps=eps)$value
  p_inv <- 1/integrate(function(x,y,sigma,eps) {inv(x,y,sigma,eps) / Z} , range[1],range[2],y,sigma,eps)$value
  integrand <- function(x){
    Bias_fun(c,x,abs=TRUE) * inv(x,y,sigma,eps) / Z
  }
  abs(integrate(integrand, range[1], range[2])$value*p_inv)
}

avg_absvar <- function(c,which_well){
  
  if (which_well =='left'){
    range <- c(-Inf,-1/sqrt(3))
  }else if (which_well == 'right'){
    range <- c(1/sqrt(3),Inf)
  }else if (which_well == 'middle'){
    range <- c(-1/sqrt(3),1/sqrt(3))
  }
  Z <- integrate(inv,-Inf,Inf,y=y,sigma=sigma,eps=eps)$value
  p_inv <- 1/integrate(function(x,y,sigma,eps) {inv(x,y,sigma,eps) / Z} , range[1],range[2],y,sigma,eps)$value
  integrand <- function(x){
    DVar_fun(c,x,abs=TRUE) * inv(x,y,sigma,eps) / Z
  }
  abs(integrate(integrand, range[1], range[2])$value*p_inv)
}
wells <- c("left","middle", "right")
c_grid <- seq(-1.5, 1.5, length.out = 3001)

# Use your integrated bias/variance functions
Bfun <- function(c, which_well) avg_absbias(c, which_well)
Vfun <- function(c, which_well) avg_absvar(c, which_well)

# -----------------------------
# Build full grid across wells
# -----------------------------
df_all <- bind_rows(lapply(wells, function(which_well) {
  Bvals <- sapply(c_grid, function(c) Bfun(c, which_well))
  Vvals <- sapply(c_grid, function(c) Vfun(c, which_well))

  data.frame(
    c = c_grid,
    absB = as.numeric(Bvals),
    absVar = as.numeric(Vvals),
    well = which_well
  )
}))

df_all <- df_all %>%
  group_by(well) %>%
  mutate(max_regret = pmax(absB, absVar)) %>%
  ungroup()

df_minimax <- df_all %>%
  group_by(well) %>%
  slice_min(order_by = max_regret, n = 1, with_ties = FALSE) %>%
  ungroup()

tol <- 0.01

lex <- df_all %>%
  group_by(well) %>%
  group_map(~{

    df <- .x
    well_val <- .y$well[[1]]

    # --------------------------------
    # Min variance, then min bias
    # --------------------------------
    Vmin <- min(df$absVar, na.rm = TRUE)

    cand_V <- df %>%
      filter(absVar <= Vmin + tol)

    if (nrow(cand_V) == 0) {
      pt_V <- df[which.min(df$absVar), ]
    } else {
      pt_V <- cand_V[which.min(cand_V$absB), ]
    }

    Bmin <- min(df$absB, na.rm = TRUE)

    cand_B <- df %>%
      filter(absB <= Bmin + tol)

    if (nrow(cand_B) == 0) {
      pt_B <- df[which.min(df$absB), ]
    } else {
      pt_B <- cand_B[which.min(cand_B$absVar), ]
    }

    data.frame(
      well = well_val,

      V_then_B_c = pt_V$c,
      V_then_B_absB = pt_V$absB,
      V_then_B_absVar = pt_V$absVar,

      B_then_V_c = pt_B$c,
      B_then_V_absB = pt_B$absB,
      B_then_V_absVar = pt_B$absVar
    )

  }) %>%
  bind_rows()

roots <- polyroot(c(-y, 1, 0, -1))
roots <- Re(roots[abs(Im(roots)) < 1e-8])

fixed_df <- bind_rows(lapply(wells, function(which_well) {
  if (length(roots) == 0) {
    data.frame(
      well = character(0),
      c = numeric(0),
      absB = numeric(0),
      absVar = numeric(0),
      type = character(0)
    )
  } else {
    data.frame(
      well = which_well,
      c = roots,
      absB = sapply(roots, function(c) Bfun(c, which_well)),
      absVar = sapply(roots, function(c) Vfun(c, which_well)),
      type = "Fixed Points"
    )
  }
}))

special_df <- bind_rows(

  # minimax
  # df_minimax %>%
  #   transmute(
  #     well,
  #     c,
  #     absB,
  #     absVar,
  #     type = "Argmin Max{E[|B|],E[|V|]}"
  #   ),

  # first minimize variance, then bias
  lex %>%
    transmute(
      well,
      c = V_then_B_c,
      absB = V_then_B_absB,
      absVar = V_then_B_absVar,
      type = "Argmin E[|B|] subj. to Argmin E[|V|]"
    ),

  # first minimize bias, then variance
  lex %>%
    transmute(
      well,
      c = B_then_V_c,
      absB = B_then_V_absB,
      absVar = B_then_V_absVar,
      type = "Argmin E[|V|] subj. to Argmin E[|B|]"
    ),

  fixed_df %>%
    transmute(
      well,
      c,
      absB,
      absVar,
      type
    )

)
special_df <- special_df %>%
  mutate(
    label_x = ifelse(well=="middle",16,50)
  )
legend_labels <- c(
  "Fixed points",
  "Minimizer of Average Variance (MoAV)",
  "Minimizer of Average Bias (MoAB)"
  #TeX('$c=\\arg\\min \\max\\{E_\\pi |B_\\theta| ,E_\\pi|\\sqrt{\\Delta V_\\theta}|\\}$')
)


p2 <- plot_average_bias_and_variance(
  df_all = df_all,
  special_df = special_df,
  legend_labels = legend_labels
)

p2
#ggsave(plot=p2,width=16,height=8,dpi=300,filename='IntegratedBiasVarianceTradeoff.png')

```
Helper functions for parameter estimation (Figure 4):
```{r warning=FALSE}
fit_for_center <- function(method,
                           sub_paths,
                           sub_times,
                           show_progress = FALSE) {

  n_paths <- length(sub_paths)
  param_labels <- c("epsilon", "y", "sigma")
  preds <- vector("list", n_paths)

  for (i in seq_along(sub_paths)) {
    if (show_progress) {
      message("fitting path ", i, " / ", n_paths)
    }
    this_path <- sub_paths[[i]]
    par0 <- c(0.5, 0, 0.1)
    obj <- function(par, path, times) {
      if (par[1] < 0.001 || par[1] > 1)
        return(1e20)
      if (par[2] < -0.37 || par[2] > 0.37)
        return(1e20)
      if (par[3] < 0.001)
        return(1e20)
      roots <- polyroot(c(-par[2], 1, 0, -1))
      real_roots <- Re(roots[abs(Im(roots)) < 1e-8])
      if (length(real_roots) < 3)
        return(1e20)
      real_roots <- sort(real_roots)
      left_fix  <- real_roots[1]
      uns_fix   <- real_roots[2]
      right_fix <- real_roots[3]
      b_vec <- NULL
      if (method == "fix") {
        b_vec <- c(left_fix, right_fix)
      } else if (method == "all fix") {
        b_vec <- c(left_fix, uns_fix, right_fix)
      } else if (method == "left fix") {
        b_vec <- c(left_fix)
      } else if (method == "wrong fix") {
        b_vec <- c(right_fix, left_fix)
      } else if (method == "random") {
        b_vec <- random_bs[i,]
      } else if (method == "average") {
        x_vals <- path[,1]
        b_vec <- mean(x_vals)
      } else if (method == "pi") {
        b_vec <- 3.14159
      } else if (method == "zero") {
        b_vec <- 0
      } else if (method == "0.5") {
        b_vec <- 0.5
      } else if (method == "uns") {
        b_vec <- uns_fix
      } else if (method == "minimax steps") {
        x_vals <- path[,1]
        c_grid <- seq(-1.5, 1.5, length.out = 2001)
        b_vec <- minimax_steps_cpp(
          x_vals, c_grid,
          par[1], par[2], par[3],
          tol=1
        )
      } else if (method == "argminBargminV steps") {
        x_vals <- path[,1]
        c_grid <- seq(-1.5, 1.5, length.out = 2001)

        b_vec <- argminBargminV_steps_cpp(
          x_vals, c_grid,
          par[1], par[2], par[3],
          tol = 1
        )
      } else if (method == "argminVargminB steps") {
        x_vals <- path[,1]
        c_grid <- seq(-1.5, 1.5, length.out = 2001)
        b_vec <- argminVargminB_steps_cpp(
          x_vals, c_grid,
          par[1], par[2], par[3],
          tol = 1
        )
      } else if (method == "minimax wells 3") {
        c_grid <- seq(-1.5, 1.5, length.out = 301)
        b_vec <- wells_bvec_trap_cpp(
          "minimax wells",
          c_grid,
          par[1], par[2], par[3],
          split_mode = "thirds",
          uns_fix,
          tol=0.2
        )
      } else if (method == "argminVargminB wells 3") {
        c_grid <- seq(-1.5, 1.5, length.out = 301)
        b_vec <- wells_bvec_trap_cpp(
          "argminVargminB wells",
          c_grid,
          par[1], par[2], par[3],
          split_mode = "thirds",
          uns_fix,
          tol = 0.2
        )
      } else if (method == "argminBargminV wells 3") {
        c_grid <- seq(-1.5, 1.5, length.out = 301)
        b_vec <- wells_bvec_trap_cpp(
          "argminBargminV wells",
          c_grid,
          par[1], par[2], par[3],
          split_mode = "thirds",
          uns_fix,
          tol = 0.2
        )
      } else if (method == "local") {
        b_vec <- path[,1]
      }
      res <- tryCatch(
        log_lik_path_rcpp(
          as.matrix(path),
          times,
          c(par[1], par[2]),
          par[3],
          method = method,
          b_vec = b_vec,
          uns_fix = uns_fix,
          fh_solver = "rk4"
        ),
        error = function(e) NULL
      )
      if (is.null(res) || !is.finite(res$ll))
        return(1e20)

      res$ll
    }

    fit <- tryCatch(
      optim(
        par = par0,
        fn = obj,
        path = this_path,
        times = sub_times,
        method = "Nelder-Mead",
        control = list(
          maxit = 5000,
          trace = 0
        )
      ),
      error = function(e) NULL
    )
    preds[[i]] <-
      if (is.null(fit) || !is.finite(fit$value))
        rep(NA_real_, length(param_labels))
      else
        fit$par
  }
  preds_mat <- do.call(rbind, preds)
  colnames(preds_mat) <- param_labels

  as.data.frame(preds_mat)
}

run_experiment <- function(
    paths,
    sub_h_vec,
    methods,
    fit_fun,
    t_0,
    N,
    h
) {

  all_estimates <- data.frame()

  for (sub_h in sub_h_vec) {

    message("Sub_h = ", sub_h)
    ratio <- sub_h/h
    sub_times <- seq(
      t_0,
      N * h,
      length.out = N / ratio + 1
    )

    idx <- seq(1, N + 1, by = ratio)

    sub_paths <- lapply(paths, function(path) {

      df <- as.data.frame(path)
      valid_idx <- idx[idx <= nrow(df)]
      df[valid_idx, , drop = FALSE]
    })

    for (k in seq_along(methods)) {

      message("  Method: ", methods[k])

      est <- fit_fun(
        method = methods[k],
        sub_paths = sub_paths,
        sub_times = sub_times
      )

      tmp <- est
      tmp$path <- seq_len(nrow(tmp))
      tmp$ratio <- ratio
      tmp$sub_h <- h * ratio
      tmp$method <- methods[k]

      all_estimates <- rbind(all_estimates, tmp)
    }
  }

  all_estimates
}
make_long_estimates <- function(all_estimates, true_params, param_labels) {

  est_long <- all_estimates %>%
    pivot_longer(
      cols = all_of(param_labels),
      names_to = "parameter",
      values_to = "estimate"
    )

  truth_df <- data.frame(
    parameter = param_labels,
    truth = true_params
  )

  est_long <- left_join(est_long, truth_df, by = "parameter")

  est_long$parameter <- factor(
    est_long$parameter,
    levels = param_labels
  )

  est_long
}
compute_metric_df <- function(est_long, metric = c("MSE", "ARE")) {

  metric <- match.arg(metric)

  est_long %>%
    mutate(
      value = if (metric == "MSE") {
        (estimate - truth)^2
      } else {
        abs(estimate - truth) / abs(truth)
      }
    ) %>%
    group_by(ratio, sub_h, method, parameter) %>%
    summarise(
      value = mean(value, na.rm = TRUE),
      .groups = "drop"
    )
}
```



## Figure 4
To generate new estimates, change make_and_save_new_estimates <- TRUE. Running new experiments requires approximately 30 minutes to simulate 1,000 paths and approximately 8 hours to estimate the parameters as described below.

If make_and_save_new_estimates <- FALSE, precomputed estimates will be used from:

read.csv("DoubleWell_ParameterEstimates/Ex1.csv")
read.csv("DoubleWell_ParameterEstimates/Ex2.csv")

```{r fig.height=8, fig.width=16, warning=FALSE}
make_and_save_new_estimates <- TRUE

if (make_and_save_new_estimates){
  M_paths <- 1000
  
  # simulate M_paths new paths of time T with true_params
  N <- 300000
  h <- 0.0001
  T <- N*h
  times <- seq(0,N*h,length.out=N+1)
  
  param_labels <- c("epsilon","y","sigma")
  true_params <- c(0.1,-0.05,1.7)
  
  #drift and diffusion
  F <- function(x,theta = true_params[1:2]){
    eps<-theta[1]; y<-theta[2]
    (x[1]-x[1]^3-y)/eps
  }
  G <- function(x, Sigma = true_params[3]) Sigma

  print("Simulating trajectories")
  paths <- EM_sim_fast(F,G,x0=0,times,M_paths,progress = TRUE)

  # run parameter estimation experiment for true_params, sub_h_vec and methods
  sub_hs <- c(0.01,0.02,0.04,0.06,0.08)
  methods <- c("random","local","wrong fix","all fix","argminVargminB steps" ,"argminBargminV steps" ,  "argminVargminB wells 3" ,"argminBargminV wells 3")
  methods <- c("random")
  
  #for random splitting only
  random_bs <- t(sapply(paths, function(path) {
    c(runif(1, min = min(min(path),-1/sqrt(3)), max = -1/sqrt(3)),
      runif(1, min = -1/sqrt(3), max = 1/sqrt(3)),
      runif(1, min = 1/sqrt(3), max = max(max(path),1/sqrt(3))))
  }))
  
  all_estimates <- run_experiment(
    paths = paths,
    sub_h_vec = sub_hs,
    methods = methods,
    fit_fun = fit_for_center,
    t_0 = 0,
    N = N,
    h = h
  )
  estimates <- make_long_estimates(
    all_estimates,
    true_params,
    param_labels
  )
  write.csv(
    estimates,
    file = "DoubleWell_ParameterEstimates/new_estimates.csv",
    row.names = FALSE
  )
}else{
  true_params_Ex1 <- c(0.1,-0.05,1.7)
  true_params_Ex2 <- c(0.25,0,1.2)

  estimates_Ex1 <- read.csv("DoubleWell_ParameterEstimates/Ex1.csv")
  estimates_Ex2 <- read.csv("DoubleWell_ParameterEstimates/Ex2.csv")
}

#choose methods to include, their colors and their labels for plots
methods <- c("random","local","wrong fix","all fix","argminVargminB steps" ,"argminBargminV steps" ,  "argminVargminB wells 3" ,"argminBargminV wells 3")

methods_keep_metric <- c(
  "all fix",
  "wrong fix",
  "random",
  "argminBargminV wells 3"
)

methods_keep_box <- c(
  "all fix",
  "wrong fix",
  "local",
  "random",
  "argminVargminB steps",
  "argminBargminV steps",
  "argminVargminB wells 3",
  "argminBargminV wells 3"
)

method_names <- c(
    "Random linearization",
    "Local Strang linearization",
    "Wrong fixed point linearization",
    "Fixed point linearization",
    "Minimizer of Local Bias (MoLB)",
    "Minimizer of Local Variance (MoLV)",
    "Minimizer of Average Bias (MoAB)",
    "Minimizer of Average Variance (MoAV)"
  )

method_colors <- c( "Random linearization" = "#1b9e77", "Local Strang linearization" = "#66a61e", "Wrong fixed point linearization" = "#1f78b4", "Fixed point linearization" = "firebrick2", "Minimizer of Local Bias (MoLB)" = "#e6ab02", "Minimizer of Local Variance (MoLV)" = "purple", "Minimizer of Average Bias (MoAB)" = "gold", "Minimizer of Average Variance (MoAV)" = "violet" )

param_labels <- c("epsilon","y","sigma")

p1 <- plot_metric_doublewell(
  estimates_Ex1,
  true_params_Ex1,
  methods,
  method_names,
  methods_keep_metric,
  method_palette=method_colors,
  title = "Experiment 1"
)

p2 <- plot_metric_doublewell(
  estimates_Ex2,
  true_params_Ex2,
  methods,
  method_names,
  methods_keep_metric,
  method_palette=method_colors,
  title = "Experiment 2"
)

mse_plot <-
  p1 /
  p2

p1 <- mse_plot +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.box = "horizontal"
  )

b1 <- plot_box_doublewell(
  estimates_Ex1,
  true_params_Ex1,
  param_labels,
  methods,
  method_names,
  method_palette = method_colors,
  ratio_value = 800,
  cap_outliers = FALSE,
  methods_keep = methods_keep_box,
  title = "Experiment 1"
)


b2 <- plot_box_doublewell(
  estimates_Ex2,
  true_params_Ex2,
  param_labels,
  methods,
  method_names,
  method_palette = method_colors,
  ratio_value = 800,
  cap_outliers = FALSE,
  methods_keep = methods_keep_box,
  title = "Experiment 2"
)

box_plot <-
  b1 /
  b2

p2 <- box_plot +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.box = "horizontal"
  )
p1
p2

ggsave(plot=p2,width=16,height=6,dpi=300,filename='DoubleWell_ParameterEstimates/RawEstimates.png')
ggsave(plot=p1,width=16,height=8,dpi=300,filename='DoubleWell_ParameterEstimates/MSE_new.png')
```